---
title: "CS770Final"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1
 The black_politicians data were collected via a field experiment where the author 
sent fictional emails purportedly sent by Black people to legislators in the United States. The experiment 
sought to determine whether the effect of the email being from "out-of-district" (someone who can’t vote 
for you and so provides no extrinsic motivation to reply) would have a smaller effect on response rates for 
Black legislators than for non-Black ones, providing evidence of additional intrinsic motivation on the 
part of Black legislators to help Black people. Implement the proper causal inference analysis for this 
dataset. 

```{r problem1}
library(boot); library(tidyverse); library(dplyr)
br <- causaldata::black_politicians

# Function to do IPW estimation with regression adjustment
ipwra <- function(br, index = 1:nrow(br)) {
    # Apply bootstrap index
    br <- br %>% slice(index)
    
    # estimate and predict propensity score
    m <- glm(leg_black ~ medianhhincom + blackpercent + leg_democrat,
             data = br, family = binomial(link = 'logit'))
    br <- br %>%
        mutate(ps = predict(m, type = 'response'))
    
    # Trim control observations outside of treated PS range
    minps <- br %>%
        filter(leg_black == 1) %>%
        pull(ps) %>%
        min(na.rm = TRUE)
    maxps <- br %>%
        filter(leg_black == 1) %>%
        pull(ps) %>%
        max(na.rm = TRUE)
    br <- br %>%
        filter(ps >= minps & ps <= maxps)
    
    # Create IPW weights
    br <- br %>%
        mutate(ipw = case_when(
        leg_black == 1 ~ 1/ps,
        leg_black == 0 ~ 1/(1-ps)))
    
    # Estimate difference
    w_means <- br %>% 
        group_by(leg_black) %>%
        summarize(m = weighted.mean(responded, w = ipw)) %>%
        arrange(leg_black)
    
    return(w_means$m[2] - w_means$m[1])
}


b <- boot(br, ipwra, R = 200)
# See estimate and standard error
b
```

## Problem 2

The close-college data is used to estimate the effect of college education on 
earnings, using the presence of a nearby (in-county) college as an instrument for college attendance. 
Implement the proper causal inference analysis for this dataset.

take example from book. Do causal inference. Use nearc4 as instrument
```{r prob2}
cc <- causaldata::close_college
# for simplicity, ignore subjects with missing outcome or missing instrument
cc.iv <- cc[which(!is.na(cc$educ) & !is.na(cc$lwage) & !is.na(cc$nearc4)),]
summary(cc.iv)

t.test(lwage ~ nearc4, data=cc.iv)

library(sem) 

model1 <- tsls(lwage ~ educ, ~ nearc4, data = cc.iv)
summary(model1)
confint(model1)  # note the wide confidence intervals
```

## Problem 3

The social-insure data is used to estimate the effect of a two-round social network-
based experiment on getting farmers to get insurance. Implement the proper causal inference analysis for 
this dataset. 


```{r prob3}
library(modelsummary); library(gmm)
d <- causaldata::social_insure
# Remove all missing observations ourselves
d <- d %>%
    select(takeup_survey, male, age, agpop, ricearea_2010,
           literacy, intensive, risk_averse, disaster_prob,
           village, address, pre_takeup_rate, default) %>%
    na.omit()

m <- gmm(takeup_survey ~ male + age + agpop + ricearea_2010 +
         literacy + intensive + risk_averse + disaster_prob +
         factor(village) + pre_takeup_rate,
         ~ male + age + agpop + ricearea_2010 +
         literacy + intensive + risk_averse + disaster_prob +
         factor(village) + default, data = d)

# We can apply the address clustering most easily in msummary
msummary(m, vcov = ~address, stars = c('*' = .1, '**' = .05, '***' = .01))
```

## Problem 4

The texas data looks at the massive expansion in prison capacity in Texas that 
occurred in 1993 under Governor Ann Richards, and the effect of that expansion on the number of Black 
men in prison. Implement the proper causal inference analysis for this dataset.

```{r prob4}
t <- causaldata::texas

library(tidyverse)
library(haven)
library(estimatr)

read_data <- function(df)
{
  full_path <- paste("https://raw.github.com/scunning1975/mixtape/master/", 
                     df, sep = "")
  df <- read_dta(full_path)
  return(df)
}

abortion <- read_data("abortion.dta") %>% 
  mutate(
    repeal = as_factor(repeal),
    year   = as_factor(year),
    fip    = as_factor(fip),
    fa     = as_factor(fa),
  )

reg <- abortion %>% 
  filter(bf15 == 1) %>% 
  lm_robust(lnr ~ repeal*year + fip + acc + ir + pi + alcohol+ crack + poverty+ income+ ur,
            data = ., weights = totpop, clusters = fip)
#weight is pct of aa population, clusters is statefip
abortion_plot <- tibble(
  sd = reg[[2]][76:90],
  mean = reg[[1]][76:90],
  year = c(1986:2000))

abortion_plot %>% 
  ggplot(aes(x = year, y = mean)) + 
  geom_rect(aes(xmin=1986, xmax=1992, ymin=-Inf, ymax=Inf), fill = "cyan", alpha = 0.01)+
  geom_point()+
  geom_text(aes(label = year), hjust=-0.002, vjust = -0.03)+
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = mean - sd*1.96, ymax = mean + sd*1.96), width = 0.2,
                position = position_dodge(0.05))
```

## Problem 5

Consider the structural nested model, $E(Y^a -Y^{a=0} | A=a, L) = \beta_1 a$. A consistent 
estimator of the casual effect of treatment $\beta_1$  can be obtained via g-estimation. In particular, the estimator 
of $\beta_1$  is the value of $H(\psi^+)=Y-\psi^+A$  that minimizes the association between $H(\psi^+)$ and  $A$. This 
value can be found through a grid search (We did it in class to find the causal effect of smoking cessation 
using the nhefs dataset. We also used the grid search values to create 95\% confidence interval for the 
causal effect of treatment by including in it all values from the grid search with p-values greater than 
0.05).

### Part A

Show that the estimator of the casual effect of treatment, 1β, can be also be obtained by solving the following estimating equation,
\[\sum_{i=1}^n I (C_i = 0)W_i^CH_i(\psi^+)[A_i - E(A|L_i)]=0\]

$$
\begin{align*}
x = 0
\end{align*}
$$

### Part B

\[\psi^+= \frac
{\sum_{i=1}^n I (C_i = 0)W_i^C Y_i[A_i - E(A|L_i)]}
{\sum_{i=1}^n I (C_i = 0)W_i^C A_i[A_i - E(A|L_i)]}\]

$$
\begin{align*}
H(\psi^+)&=Y-\psi^+A\\
H(\psi^+)&=Y- 
\frac{\sum_{i=1}^n I (C_i = 0)W_i^C Y_i[A_i - E(A|L_i)]}{\sum_{i=1}^n I (C_i = 0)W_i^C A_i[A_i - E(A|L_i)]}A
\end{align*}
$$